---
output: 
  github_document:
    df_print: tibble
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# easyphreeqc

The goal of easyphreeqc is to provide a more useful interface to the existing phreeqc package.

## Installation

You can install easyphreeqc from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("paleolimbot/easyphreeqc")
```

## Example

Running PHREEQC is accomplished using the `phreeqc()` function, which calls the program and generates the output. The function accepts character vectors of input, which can be generated using intput helper functions such as `solution()`, `selected_output()`, `equilibrium_phases()`, and `reaction_temperature()` (or roll your own input using `phreeqc_input()`). To get the raw output of PHREEQC, you will have to pass `quiet = FALSE`.

```{r}
library(easyphreeqc)
phreeqc(
  solution(pH = 7, temp = 25),
  quiet = FALSE
)
```

To get the results as a data frame, we need to supply a `selected_output()` to the input file.

```{r}
phreeqc(
  solution(pH = 7, temp = 25),
  selected_output(pH = TRUE, temp = TRUE, activities = c("OH-", "H+", "O2"))
)
```

To find the distribution of a few solutions, you can add solutions to the input (you will have to number them).

```{r}
phreeqc(
  solution(1, pH = 6, temp = 25),
  solution(2, pH = 7, temp = 25),
  solution(3, pH = 8, temp = 25),
  selected_output(pH = TRUE, temp = TRUE, activities = c("OH-", "H+", "O2"))
)
```

## Databases

Some elements (for example, mercury) aren't included in the base database. There are a number of databases included in the PHREEQC package, that you can choose by specifying the `db` argument of `phreeqc()`. One that includes mercury is the "minteq" database.

```{r}
phreeqc(
  solution(1, pH = 7, temp = 25, Hg = 0.1),
  solution(2, pH = 7, temp = 25, Hg = 0.2),
  solution(3, pH = 7, temp = 25, Hg = 0.3),
  selected_output(activities = c("Hg", "Hg2+2", "Hg(OH)2", "Hg(OH)2",
                                 "HgOH+", "Hg(OH)3-")),
  db = "minteq"
)
```

(Note that I called `phreeqc()` with `quiet = FALSE` to first find the species included in `selected_output()`)

## Multiple solutions

The input for `phreeqc()` can accept `list()` objects, so you can use something like `plyr::mlply()` to create a `list` of solutions like the one above to pass to `phreeqc()`. You could use `expand.grid()` to generate a large number of solutions in this way.

```{r}
# create data.frame with arguments to solution()
solution_info <- data.frame(
  .number = 1:3,
  pH = 7,
  temp = 25,
  Hg = c(0.1, 0.2, 0.3)
)

# use mlply to turn this into a list()
solution_list <- plyr::mlply(solution_info, solution)

# pass to phreeqc()
phreeqc(
  solution_list,
  selected_output(activities = c("Hg", "Hg2+2", "Hg(OH)2", "Hg(OH)2",
                                 "HgOH+", "Hg(OH)3-")),
  db = "minteq"
)
```
